Program Transformations for miniKanren

### 2020-05-22 papers/miniKanren/text/nonconj.tex

@@

In this section we will describe a novel approach to specialization of relational programs.
This approach draws inspiration from both conjunctive partial deduction and supercompilation.
The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
\db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
    \db{For the sake of brevity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
      Driving itself is a trivial combination of presented functions (line 2)}}}.

@@

Unfolding is a process of substitution of some relation call by its body with simultaneous computation of unifications.
To unfold a relation call we do the following steps.
First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
All fresh variables get instantiated.
The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
All unifications are computed.
Those disjuncts in which unifications fails are removed.
Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.

>>>>

\newcommand{\code}[1]{\texttt{#1}}

@@

In this section we will describe a novel approach to specialization of relational programs.
This approach draws inspiration from both conjunctive partial deduction and supercompilation.
The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
\db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
    \db{For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
      Driving itself is a trivial combination of presented functions (line 2).}}}.

\db{
Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
Moreover, we assume all fresh variables to be instantiated and all unifications to be computed on each step.
Those disjuncts in which unifications fails are removed.
Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
Any miniKanren term can be trivially transformed to the described form\footnote{
  \db{On Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization;
    The code is ommited for brevity.}
}.}

@@

Unfolding is a process of substitution of some relation call by its body with simultaneous computation of unifications.
To unfold a relation call we do the following steps.
\db{TODO: the following has alredy mentioned above.}
First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
All fresh variables get instantiated.
The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
All unifications are computed.
Those disjuncts in which unifications fails are removed.
Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
